---
title: "Data 607 - Project 1"
author: "William Forero"
output:
  html_document: default
  pdf_document: default
---

## Project Objective

In this project, you’re given a text file with chess tournament results where the information has some structure. Your
job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database) with the following information for all of the players:

Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents

For the first player, the information would be:
Gary Hua, ON, 6.0, 1794, 1605

1605 was calculated by using the pre-tournament opponents’ ratings of 1436, 1563, 1600, 1610, 1649, 1663, 1716, and
dividing by the total number of games played.

The chess rating system (invented by a Minnesota statistician named Arpad Elo) has been used in many other contexts,
including assessing relative strength of employment candidates by human resource departments.

```{r}
library(tidyverse)
library(stringr)
library(dplyr)
```


## Importing The Data

First we load the raw data from the text file and look at how it is structured.
```{r}
raw_tournament_data <- readLines("https://raw.githubusercontent.com/william-forero/Data-607/refs/heads/main/tournamentinfo.txt")

head(raw_tournament_data,21)
```
We can see that there is a repetitive structure. 

Each players information is split into two rows:
Row 1 always begins with a number 
Row 2 always begins with a letter

Players are split by a row of "---"

## Parsing The Data

First I will remove the headers and any rows that are all '---'
```{r}
# Removing the lines consisting of all hyphens
tournament_data <- raw_tournament_data[!str_detect(raw_tournament_data, "^-+$")]

# Remove the headers
tournament_data <- tournament_data[-(1:2)]

head(tournament_data,10)
```
Next we parse out the two player row formats
```{r}
# Checking the first 7 characters of the row and determining if it has a number
player_row_1 <- tournament_data[str_detect(str_sub(tournament_data, 1, 7), "[0-9]")]

# Checking the first 7 characters of the row and determining if it has a letter
player_row_2 <- tournament_data[str_detect(str_sub(tournament_data, 1, 7), "[A-Z]")]

head(player_row_1)
head(player_row_2)
```
Now we break each field out into its own column. The format will depend on the row type (row 1 or 2 of the players data)
```{r}
# First I split by the delimeter |
player_row1_df <- read_delim(I(player_row_1),"|",trim_ws = TRUE, col_names = FALSE)

# Rename the columns
names(player_row1_df) <- c("pair_num","player_name","total_points",
                           "round1","round2","round3","round4","round5",
                           "round6","round7")

player_row1_df <- player_row1_df |> select(-last_col())
head(player_row1_df)

```

Next we parse out the rows of the second format
```{r}
# Parsing out the 2nd row of each players data
player_row2_df <- data_frame(player_state = str_trim(substr(player_row_2, 1,  6), side="both"),
                             uscf_id = substr(player_row_2, 8,  16),
                             pre_rating = as.numeric(substr(player_row_2, 22, 26)),
                             post_rating = as.numeric(substr(player_row_2, 32, 35))
                             )
                             

head(player_row2_df)
```

Now we combine the two dfs and pivot longer
```{r}
# We combine the two data frames into one comprehensive player data frame
full_player_data <- bind_cols(player_row1_df,player_row2_df)

# Now we pivot longer so that the rounds become rows, we also split outcome and opponent
player_data_long <- full_player_data |>
  pivot_longer(
    cols = starts_with("round"),
    names_to = "round",
    values_to = "round_result"
  ) |>
  extract(
    round_result, 
    into = c("rd_outcome", "rd_opponent"), 
    regex = "([A-Z])\\s*(\\d+)"
  ) |>
  # This is used to remove round rows where the player did not have an opponent
  filter(!is.na(rd_opponent)) |>
  mutate(rd_opponent = as.numeric(rd_opponent))


head(player_data_long)

```

We can create a column that provides the opponents pre-rating by first creating an intermediary table that contains the player id as well as the pre_rating.
Once we have that we can join the opponents pre_rating onto the larger df to be used to calculate the average opponent pre-rating.
```{r}
# Creating an intermediary table to pull the opponents rating
player_row1_subset <- player_row1_df |> select(1)
player_row2_subset <- player_row2_df |> select(3)
player_pre_ratings <- bind_cols(player_row1_subset,player_row2_subset)

player_data_long <- player_data_long |>
  left_join(player_pre_ratings, by = c("rd_opponent" = "pair_num")) |>
  rename(`opp_pre_rating` = pre_rating.y) |>
  rename(`pre_rating` = pre_rating.x)

head(player_data_long)
```

Now we can extract the fields we want to provide in csv format
```{r}
# Calulate the average opponent rating
final_player_results <- player_data_long |>
group_by(pair_num,player_name, player_state, total_points, pre_rating) |>
summarise(avg_opp_pre_rating = round(mean(opp_pre_rating, na.rm = TRUE)))

print(final_player_results)
```

## Export to CSV

Now that we have the data in the format that we want, we can export to a CSV file.
```{r}
write.csv(final_player_results, "tournamentinfo.csv")
```






